Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-09-10 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2: Binding notation
Fix typo

The aim of this paper is to present detailed formalizations of variants of ontological arguments in a natural deduction calculus. For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz} or \citep{gabbay}. The natural deduction calculus proposed and used in this paper is an extension of a standard natural deduction calculus for (non-modal) higher-order logic \citep{BezmuellerBrownKohlhase} with new introduction and elimination rules for modal operators (as presented in Section \ref{sec:calculus}). This calculus is sound and complete relative to the non-modal calculus extended with Axiom K and the Necessitation rule (as demonstrated in Subsection \ref{sec:compl-sound}). There are various higher-order logics \citep{BezmuellerBrownKohlhase} and various higher-order modal logics \citep{Muskens}. The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles. With additional axioms for stronger modal logics, such as KB and S5, this calculus is sufficient for achieving the stated aim.

@@

\begin{definition} \emph{Terms and formulas} are given by the following grammar:
\begin{align*}
 s,t \quad ::= \quad & 
  p_\tau \ \mid \ 
  X_\tau \ \mid \
  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
& \bot_o \ \mid \
  \imp_{o\imp o\imp o} \ \mid \ 
  \wedge_{o\imp o\imp o} \ \mid \\
& \all_{(\tau\imp o)\imp o} \ \mid \ 
  \ex_{(\tau\imp o)\imp o} \ \mid \
  \nec_{o\imp o} \ \mid \
  \pos_{o\imp o}
\end{align*}
where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$. Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed. Furthermore, subscript types are omitted when they are clear from the context. Implicit alpha-conversion is assumed: two expressions are considered to be the same modulo renaming of their bound variables.
Negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$), equivalence ($\biimp_{o\imp o\imp o}$), equality ($=_{\tau\imp\tau\imp o}$), inequality ($\neq_{\tau\imp\tau\imp o}$), are defined by $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$, $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$, $s = t \equiv \forall P. (P \ s) \biimp (P \ t)$ (Leibniz's equality) and $s \neq t\equiv \neg (s=t)$, .
\end{definition}

>>>>

The aim of this paper is to present detailed formalizations of variants of ontological arguments in a natural deduction calculus. For a comprehensive introduction to natural deduction, the reader can consult \citep{prawitz} or \citep{gabbay}. The natural deduction calculus proposed and used in this paper is an extension of a standard natural deduction calculus for (non-modal) higher-order logic \citep{BezmuellerBrownKohlhase} with new introduction and elimination rules for modal operators (as presented in Section \ref{sec:calculus}). This calculus is sound and complete relative to the non-modal calculus extended with Axiom K and the Necessitation rule (as demonstrated in Subsection \ref{sec:compl-sound}). There are various higher-order logics \citep{BezmuellerBrownKohlhase} and various higher-order modal logics \citep{Muskens}. The calculus proposed here can be regarded as a minimalistic calculus for a rigid higher-order modal logic K without extensionality principles. With additional axioms for stronger modal logics, such as KB and S5, this calculus is sufficient to achieve the stated aim.

@@

\begin{definition} \emph{Terms and formulas} are given by the following grammar:
\begin{align*}
 s,t \quad ::= \quad & 
  p_\tau \ \mid \ 
  X_\tau \ \mid \
  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
& \bot_o \ \mid \
  \imp_{o\imp o\imp o} \ \mid \ 
  \wedge_{o\imp o\imp o} \ \mid \\
& \all_{(\tau\imp o)\imp o} \ \mid \ 
  \ex_{(\tau\imp o)\imp o} \ \mid \
  \nec_{o\imp o} \ \mid \
  \pos_{o\imp o}
\end{align*}
where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$. Parenthesis conventions, infix notation for propositional connectives and binding notation\footnote{$\forall x. F[x]$ and $\exists x. F[x]$ are assumed to be abbreviations for, respectively, $\forall (\lambda x. F[x])$ and $\exists (\lambda x. F[x])$.} for quantifiers are assumed. Furthermore, subscript types are omitted when they are clear from the context. Implicit alpha-conversion is assumed: two expressions are considered to be the same modulo renaming of their bound variables.
Negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$), equivalence ($\biimp_{o\imp o\imp o}$), equality ($=_{\tau\imp\tau\imp o}$), inequality ($\neq_{\tau\imp\tau\imp o}$), are defined by $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$, $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$, $s = t \equiv \forall P. (P \ s) \biimp (P \ t)$ (Leibniz's equality) and $s \neq t\equiv \neg (s=t)$, .
\end{definition}
